---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# funnelplot

  <!-- badges: start -->
  [![Travis build status](https://travis-ci.org/oizin/funnelplot.svg?branch=master)](https://travis-ci.org/oizin/funnelplot)
  [![codecov](https://codecov.io/github/oizin/funnelplot/branch/master/graphs/badge.svg)](https://codecov.io/github/oizin/funnelplot)
  <!-- badges: end -->
  
  
The goal of funnelplot is to simplify risk adjusted comparisons of institutional performance for the purpose of outlier discovery.

## Installation

You can install the development version of funnelplot from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("oizin/funnelplot")
```
## Example

This is a basic example which shows you how to produce a funnel plot:

```{r example}
library(funnelplot)
data("example_data")

# outcome ~ covariates | cluster ID
f1 <- funnel(test ~ 1 | hosp_id, data = example_data)
plot(f1)
```

See the vignettes:

```{r,eval=FALSE}
# vignettes
vignette("funnelplot-intro")
```

## References

Benjamini, Y., & Hochberg, Y. (1995). Controlling the false discovery rate: a practical and powerful approach to multiple testing. Journal of the royal statistical society. Series B (Methodological), 289-300.

Jones, H. E., Ohlssen, D. I., & Spiegelhalter, D. J. (2008). Use of the false discovery rate when comparing multiple health care providers. Journal of clinical epidemiology, 61(3), 232-240.

Spiegelhalter, D. J. (2005). Funnel plots for comparing institutional performance. Statistics in medicine, 24(8), 1185-1202.
