---
title: "Introduction to funnelplot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro-funnelplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 5,
  fig.width = 8
)
```

Funnel plots are designed for comparisons of institutional performance (e.g. hospital mortality rates or store sales), for the purpose of outlier discovery. They graph institution performance y (on a metric of interest) against institution size n. Control limits are then overlayed, designed to capture (1-$\alpha$)100% of institutions (typically 95% and 99.7% - 2 and 3 standard deviations), from a target value (e.g. the global average) for the metric. The "funnel" shape comes from the expectation that with increasing institution size there will be less variation around the target, leading to narrower control limits as n grows. For an excellent introduction see Spiegelhalter (2005).

```{r setup}
library(funnelplot)
library(ggplot2)
```

To illustrate the funnelplot package, we will use a synthetic example dataset of 40 hospitals. The binary outcome measure is whether a particular procedure was successful on the ith patient. We have two variables gender and age that may be distributed unevenly across the hospitals and play a role in explaining differences in the observed hospital success rate y = sum test / n.

```{r}
data("example_data", package = "funnelplot")
head(example_data)
```

### Unadjusted funnel plot 

We first create an unadjusted funnel plot. The resulting plot shows the raw event proportions against the number of observations per cluster. There appear to be some outliers.

```{r}
# outcome ~ covariates | cluster ID
f1 <- funnel(test ~ 1 | hosp_id, data = example_data)
plot(f1) +
  theme_funnel()
```

The plot returned by funnelplot::plot is a standard ggplot2 object, allowing a large degree of customisation to fit the end use.  

```{r}
pp <- plot(f1)
class(pp)
pp +
  theme_bw()
```

funnelplot contains other useful functions such as `outliers` that returns ... and a `summary` method that ....

```{r}
outliers(f1)
```

### Risk adjustment 

Generally raw comparison of event rates between institutions runs the risk of failing to account for differences in the populations between each cluster (e.g. a tendency for sicker patients to attend a particular hospital). To account for this a statistical model can be used to estimated hat{y} for each observation in the data - the predicted outcome given the observed covariates - which can be compared to the observed outcome y to obtain standardised rates per institution. Currently funnelplot uses logistic regression `stats::glm` for risk adjustment. A standard R formula of the form `outcome ~ covariates | clusterID` creates a risk adjusted funnel plot object. 

```{r}
f2 <- funnel(test ~ gender | hosp_id, data = example_data)
plot(f2)
```

### Adjusting for multiplicity and overdispersion

funnelplot can ... through the `pointTarget` function, that is passed to the control argument. Using `pointTarget` we can also pass an arbitrary number of control limits to funnel. In comparison to the above plots we now see that ... 

```{r}
f3 <- funnel(test ~ gender | hosp_id, data = example_data)
plot(f3)
```

### Standardised rates

By default funnelplot produces risk adjusted rates, but in many circumstances standardised rates (of the form observed/expected) will be of more interest.

```{r}
f4 <- funnel(test ~ gender | hosp_id, data = example_data)
plot(f4)
```

### Plotting

Using the label and identify arguments we can label or plot only a selection of hospitals if desired. These arguments also accept "outliers" as a value. 

```{r}
plot(f4,label="outliers")
```

```{r}
plot(f4,identify=13,label=13)
```


### References

Benjamini, Y., & Hochberg, Y. (1995). Controlling the false discovery rate: a practical and powerful approach to multiple testing. Journal of the royal statistical society. Series B (Methodological), 289-300.

Jones, H. E., Ohlssen, D. I., & Spiegelhalter, D. J. (2008). Use of the false discovery rate when comparing multiple health care providers. Journal of clinical epidemiology, 61(3), 232-240.

Spiegelhalter, D. J. (2005). Funnel plots for comparing institutional performance. Statistics in medicine, 24(8), 1185-1202.
